Multiplex stream graphs

### 2019-05-30 tex/main.tex
remise en forme de main

	\subsection{Definition}
    We define a multiplex stream graph as a tuple $M = (T,T_M,V,W_M,E_M,{\cal L})$, such that $T$ and $V$ are, respectively, a time interval and a set of nodes, just like for stream graphs \cite{stream}.
    As for multilayer graphs \cite{mlkiv}, ${\cal L}$ is a set of $d$ {\em aspects}; ${\cal L} = \{L_i\}_{i=1}^d$, and an element $\alpha_i$ of $L_i$ is named {\em elementary layer}. An element of $L=L_1\times \dots \times L_d$ is named {\em layer}.

	$T_M$ is a set of intervals of $T$ so for that each layer $\alpha$, $T_{\alpha}$ is the interval at which the layer exists. For each $t$ in $T$, $\exists \alpha \in L | t \in T_{\alpha}$ (at any time, all the aspects are non-empty).

	$W_M \subseteq T\times V \times (L_1 \times \dots \times L_d)$ represents the existing points on layers with respect to time.

   $E_M \subseteq T\times V_M \otimes V_M$ are the links appearing with respect to time,  $V_M = \{ (u,\alpha), u\in V, \alpha \in L\}$. The elements of $V_M$ are named {\em node-layer}.

    Notice that just like in multiplex networks, nodes in time ({\em i.e.}, elements of $W_M$) can be present in an arbitrary number of layers.

@@

	\subsection{2 examples}
	\begin{ex}[Research world]\label{exresearch}
	We consider a set of people working in research.

@@

	The multilayer stream graph is the following :
	\begin{itemize}
		\item the {\bf nodes V} are the researchers
		\item the {\bf aspects L} are : the university/companies , the departments, the subjects, the type of status of interaction (supervisor/ supervised / collaborator) 
		\item {\bf links} appear during an interaction between two researchers.
		\item all these parameters vary with time, the set $T$ is chosen arbitrarily
	\end{itemize}

	Question : which aspect is the more important in the spreading and creation of knowledge ?

@@

	\begin{figure}
    	\centering
    	\includegraphics[width=15cm]{schemas/chercheurs.jpg}
    	\caption{An simplified scheme of a multiplex stream graph for example \ref{exresearch}. The nodes are the researchers R1,R2,R3, and the aspects are the Universities (U1,U2,U3), and the department (Maths, Computer sciences, Physics)}
    	\label{fig:chercheurs}
	\end{figure}

@@

		The stream-multilayer graph is the following : 
		\begin{itemize}
			\item the {\bf nodes V} are the individuals
			\item the {\bf aspects L} are : the species, the type of possible interaction, the state of the individual (sane, infected, contagious, dead, recoved ... ) 
			\item {\bf links} appear when there is an interaction between two individuals.
			\item all these parameters vary with time, the set $T$ is chosen arbitrarily (from example from the start of the infection)
		\end{itemize}

@@

   	\subsection{Substreams and operations over substreams}

    From this definition, let us define common streams of interest.

	For any layer $\alpha \in L_1 \times \dots \times L_d$, the {\em intra-layer} stream graph $S^{\alpha}$ is the stream $S^{\alpha}=(T^{\alpha},V^{\alpha},W^{\alpha},E^{\alpha})$ such that for all $t$ in T :
	\begin{itemize}
		\item $T^{\alpha} = T_{\alpha} \in T_M$
		\item $V^{\alpha} =\{(u,\alpha),u\in V\} \cap V_M$
		\item $W^{\alpha} =\{(t,u,\alpha),|u\in V, t \in T \} \cap W_M$
		\item $E^{\alpha} = \{(t,(u,\alpha),(v,\alpha))| t \in T,(u,v)\in V^2 \} \cap E_M$.
	\end{itemize}

	For any two layers $\alpha, \beta \in L_1\times \dots\times L_d$, the {\em inter-layer} stream graph is the stream $S^{(\alpha,\beta)} = (T, V^{\alpha,\beta},W^{\alpha,\beta},E^{\alpha,\beta})$ such that :
	\begin{itemize}
		\item $T^{\alpha,\beta}=T^{\alpha}\cap T^{\beta}$
		\item $V^{\alpha,\beta} = \{u,\gamma | u \in V, \gamma \in \{\alpha,\beta\} \} \cap V_M$
		\item $W^{\alpha,\beta}= \{(t,u,\gamma) | u \in V, \gamma \in \{\alpha,\beta\} \} \cap W_M$
	    \item $E^{\alpha,\beta} = \{(t,(u,\alpha),(v,\beta) | u,v \in V\} \cap E_M $
	\end{itemize}

	The {\em intersection} of two substreams $S_1$ and $S_2$ is defined as follows :
	\[
		S' = S_1 \cap S_2 = (T_1\cap T_2, V_1 \cap V_2, W_1 \cap W_2, E_1\cap E_2)
	\]

	The {\em union} of two substreams $S_1$ and $S_2$ is defined as follows :
	\begin{align*}
		S' = S_1 \cup S_2 = (T', V_1 \cup V_2, W_1 \cup W_2, E' \})\\
		T' = [\min(T_1,T_2),\max(T_1,T_2)]\\
		E' = E_1 \cup E_2 
	\end{align*}

	Notice that the union of two sub stream graphs gives a stream graph.

	\subsection{Projections }

@@

	\begin{figure}
		\centering
		\includegraphics[width=\textwidth]{schemas/edgesCat.jpg}
		\caption{Illustration with our example of coupling edges, intra-layer edges and inter-layer edges}
		\label{exIntraInter}
	\end{figure}
   	The {\em multilayer graph at time t} $M_t$ is $M_t = (V_{M,t}, E_{M,t},V_t,{\cal L}_t)$ with :
    \begin{itemize}
		\item $V_{M,t} = \{(u,\alpha)| (t,u,\alpha)\in W_M\}$
		\item $E_{M,t} = \{(u,\alpha,v,\beta) | (t,u,\alpha,v,\beta) \in E_M\}$
		\item $V_t = {u | (t,u) \in V}$
		\item ${\cal L}_t = {L_i,t}_{i=1}^d, L_{i,t}=(\alpha)_i, t\in T_\alpha$
    \end{itemize}

	\begin{figure}	
		\begin{minipage}{0.49\linewidth}
			\includegraphics[width=\textwidth]{schemas/pauset.jpg}
		\end{minipage}
		\begin{minipage}{0.49\linewidth}
			\includegraphics[width=\textwidth]{schemas/pausetproj.jpg}
		\end{minipage}
		\caption{Our multilayer example at time $t=5$}
	\end{figure}

    \begin{rmq}
    	We assume that $\forall (i,t), \; L_{(i,t)} \neq \emptyset$.
    \end{rmq}

    A {\em multiplex stream graph (MpSG)} is a graph in which the only inter-layer connections allowed are connections between the same point : $E_M = E_I \cup E_C$.

    The {\em multilayer graph induced} $M_I(M)$ of $M$ is the multilayer graph which gather all the points on all the layer who appear during $T$ and all the links in $E$.
    \begin{align*}
    	M_I(M) = (V_{M,I}, E_{M,I}, V_I,L_I)\\
    	V_{M,I} = \bigcup_{t\in T} V_{M,t}\\
    	E_{M,I} = \bigcup_{t\in T} E_{M,t}\\
    	V_I = \bigcup_{t\in T} V_t \\
    	L_I = \bigcup_{t\in T} L_t\\
    \end{align*}

    The {\em underlying stream } $S_U(M)$ of $M$ is $(T,V_M,W_M,E_M)$, and could be share into clusters corresponding to the different layers. This definition is a generalisation to the one used for simple multilayer graphs.

	\begin{prop}
		\[
			\bigcup_{(\alpha,\beta) \in L^2} S^{(\alpha,\beta)} = S_U(M)
		\]
	\end{prop}
	\begin{proof}
		\begin{align*}
			S=\bigcup_{(\alpha,\beta) \in L^2} S^{(\alpha,\beta)}\\
			S=(T^{*},V^{*},W^{*},E^{*})\\
			S_U(M) = (T_U,V_U,W_U,E_U)			
		\end{align*}

	\begin{itemize}
		\item $T=T_U$ by definition of $S_U$. Let's take $t\in T$. By definition of multilayer stream graphs, $\exists \alpha \in L | t\in T_{\alpha}=T^{\alpha,\alpha}$. So $t \in [\underset{(\alpha,\beta) \in L^2}{\min}(T^{\alpha,\beta}),\underset{(\alpha,\beta) \in L^2}{\max}(T^{\alpha,\beta})]=T^{*}$. $T_U \subseteq T^{*}$. 
	Moreover, $\underset{(\alpha,\beta) \in L^2}{\min}(T^{\alpha,\beta}) \geq \min(T)$ 
	and $\underset{(\alpha,\beta) \in L^2}{\max}(T^{\alpha,\beta}) \leq \max(T)$. 
	So $\mathbf{T^{*} = T}$.
		\item  $V_M = V_U$ by definition of $V_U$.$V^{*}=\bigcup_{(\alpha,\beta) \in L^2} V^{\alpha,\beta} \subseteq V_M$.
	Let $(v,\alpha) \in V_U$. $(v,\alpha) \in V^{\alpha,\alpha} \in V^{*}$.$V^{*}=V_U$.$\mathbf{V^{*}=V_U}$.
		\item $W_U=W_M$ and $W^{*}=\bigcup_{\alpha,\beta \in L^2} W^{\alpha,\beta} \subset W_M$. So as for all $ (t,u,\alpha)$ in $W_M$, $(t,u,\alpha)$ belongs to $W^{\alpha,\alpha}$, $W_M$ is a subset of $W^{*}$. So $\mathbf{W^{*}=W_M}$.
		\item $E_U=E_M$ and $E^{*}=\bigcup_{\alpha,\beta \in L^2}(E^{\alpha,\beta})$ by definition. So as $E^{\alpha,\beta}$ is a subset of $E_M$ for every couple of layers $\alpha,\beta$, $E^{*} \subset E_M$. 
		$\forall e =(t,u,\alpha,v,\beta) \in E_M$, $e \in E^{\alpha,\beta}$. So $\mathbf{E_M=E^{*}}$.
	\end{itemize}		
	\end{proof}

@@

	 The {\em contribution} of one node-layer in the MPG is $n_{v,\alpha} = \frac{|T_{v,\alpha}|}{|T|}$. The {\em number of node-layers} is the sum of the contributions $N(M) = \underset{(u,\alpha)\in V_M}{\sum} n_{(u,\alpha)} = \frac{|W_M|}{|T|}$.

**************NEW*******************

    We define too the {\em contribution} of one layer as follows : $n_\alpha = \frac{|T_{\alpha}|}{|T|}$. The {\em number of layers is then the sum of the contributions}.

	In multilayer graphs, the number of nodes hasn't been fully defined as far as we know.
	Rather than just counting the nodes or the node layers, we define the {\em number of nodes in a multilayer} as :
	\begin{align*}
		N_{nodes in multilayers}(M_l) = \frac{\text{number of node-layers}}{\text{number of layers}}
	\end{align*}	    
	Notice that if there is only one layer, we find the classical number of nodes. A node appearing in half of the layers doesn't bring the same contribution than a node appearing in all the layers.

@@

    The {\em the contribution} of a node describes the rate of appearance of the node on the different layers : $n_v = \frac{\sum_{\alpha \in L}|T_{v,\alpha}|}{\sum_{\alpha \in L} |T_{\alpha}|}$.
    The {\em number of nodes} is then the sum of the contribution of the nodes : 
    \begin{align}
    N_n(M) = \sum_{v\in V} n_v= \sum_{v\in V} \frac{\sum_{\alpha \in L}|T_{v,\alpha}|}{\sum_{\alpha \in L} |T_{\alpha}|} 
    \label{numberNodes}
	\end{align}     
**********END nEW*******************

	The {\em overlap presence } of two nodes-layers $(u,\alpha)$ and $(v,\beta)$ measures the rate of simultaneous presence of two node-layers :
	\begin{align*}
		\mathbb{U}((u,\alpha),(v,\beta))=\mathbb{P}( t \in T_{u,\alpha} \cap T_{v,\beta} | t \in T_{u,\alpha} \cup T_{u,\beta}) \\
		= \frac{|T_{u,\alpha}\cap T_{u,\beta}|}{|T_{u,\alpha}\cup T_{u,\beta}|}
	\end{align*}

	The {\em overlap presence} of two layers $\alpha$ and $\beta$ is defined on the same pattern :
	\begin{align*}
		\mathbb{U}(\alpha,\beta) = \frac{|T_{\alpha}\cap T_{\beta}|}{|T_{\alpha}\cup T_{\beta}|}
	\end{align*}

    The {\em node-layer uniformity} measures the rate of simultaneous presences of node-layers in the multilayer-stream graph :
    \[
    	\mathbb{U}(M) = \frac{\sum_{(u,\alpha),(v,\beta) \ V_M \otimes V_M}{|T_{(u,\alpha)} \cap T_{(v,\beta)}|}}{\sum_{(u,\alpha),(v,\beta) \ V_M \otimes V_M}{|T_{(u,\alpha)}\cup T_{(v,\beta)}|}}
    \]

@@

	When $\mathbb{U}=1$, we say that the multilayer stream graph is uniform.

>>>>

	\subsection{Preliminaries}

	\pimprenelle{ajout des définitions préliminaires. J'ai essayé d'être concise, tout en renvoyant à nos deux papiers de référence }

	To understand well the notion of multilayer stream graph, it is important to know the multilayer graphs and the stream graphs. We will give a short definition of each object.

	\begin{defn}[Graph]
	    A graph $G=(V,E)$ is a set of "vertices" V, and of "edges" E. An edge represents an interaction between two vertices.
	\end{defn}

	\begin{defn} [Multilayer graph]
    A multilayer graph \cite{mlkiv} is a 4-element set $M=(V_M,E_M,V,L)$. V is a set of nodes which can belong to different layers. The structure of the layers is modeled by $L=(L_1,\dots L_n)$. Each $L_i$ is a set of "elementary layers". An elementary layer is an n-uplet of ${\bf L} = L_1\times L_2 \times \dots L_n$. A node in a layer is called "node-layer", the corresponding set is $V_M$. Then $E_M$ is the set of edges between node-layers.
	\end{defn}

	The multilayer graphs have been created to describe graphs in which d nodes  can have different properties (represented by the layers) and diff

	\begin{defn}[Stream graph]
	   \cite{stream} Let $T$ be a time interval, and $V$ a set of nodes. $W$ is a set of elements $(t,v), t\in T$ and represents the apparition of nodes during the time. $E$ is a set of triplets $(t,u,v)$ such as $(t,u)$ and $(t,v)$ are in $W$, and represents the links as function of time. A {\em stream graph} is the quadruplet $S=(T,W,V,E)$.

	    We call $T_u$ the set of time $t$ such as $(t,u)$ is in W, and $T_{uv}$ contains all the $t$ such as $(t,u,v)$ are in $E$.
	\end{defn}

    Through this paper, each time we will define a new "measure" for multilayer stream graphs, we will make the link with the corresponding notion in the multilayer graphs and in the stream graphs, when it is relevant. This will demonstrate that multilayer stream graphs are a generalisation of the two concepts.

	\subsection{Definition}	
	\begin{defn}[Multilayer stream graph]
    We define a multilayer stream graph as a tuple $S_M = (T,T_M,V,W_M,E_M,{\cal L})$, such that $T$ and $V$ are, respectively, a time interval and a set of nodes, just like for stream graphs \cite{stream}.

    As for multilayer graphs \cite{mlkiv}, ${\cal L}$ is a set of $d$ {\em aspects}; ${\cal L} = \{L_i\}_{i=1}^d$, and an element $\alpha_i$ of $L_i$ is named {\em elementary layer}. An element of $L=L_1\times \dots \times L_d$ is named {\em layer}.

	$T_M$ is a set of intervals of $T$ so for that each layer $\alpha$, $T_{\alpha}$ is the interval at which the layer exists. For each $t$ in $T$, $\exists \alpha \in L | t \in T_{\alpha}$. At each time of $T$, at least one layer of $L$ exists.

	$W_M \subseteq T\times V \times (L_1 \times \dots \times L_d)$ represents the existing points on layers with respect to time. For each element $(t,u,\alpha)$ of $W_M$, $t$ must be in the interval of existence of $\alpha$.

   $E_M \subseteq T\times V_M \otimes V_M$ are the links appearing with respect to time,  $V_M = \{ (u,\alpha), u\in V, \alpha \in L\}$. The elements of $V_M$ are named {\em node-layer}. A link cannot appear outside the time of existence of the two node-layers.

    Notice that just like in multiplex networks, nodes in time ({\em i.e.}, elements of $W_M$) can be present in an arbitrary number of layers.
	\end{defn}

@@

   	\pimprenelle{Ajout des contraintes sur les temps d'existence des node-layers et des links}

	\subsection{2 examples}
	\begin{ex}[Research world]\label{exresearch}
	We consider a set of researchers.

@@

	In this multilayer stream graph, the {\bf nodes V} are the researchers. The {\bf aspects L} are the university/companies , the departments, the subjects, the type of status of interaction (supervisor/ supervised / collaborator). {\bf Links} appear during an interaction between two researchers, for example if they work together on the same paper. The layers, researchers, and collaborations vary with time, this gives us the stream aspect.

@@

	\begin{figure}[h]
    	\centering
    	\includegraphics[width=15cm]{schemas/chercheurs.jpg}
    	\caption{An simplified scheme of a multiplex stream graph for example \ref{exresearch}.\newline
    	$T=[0,10], V=(R_1,R_2,R_3), L = [[U1,U2,U3],[Math,Info,Physics]]$\newline
    	$T_M=([0,10]\times\{[U1,Maths],[U1,Info]\})\cup([4,10]\times\{[U2,Maths]\})\cup\dots$\newline
    	$W_M=[0,10]\times \{(R1,[U1,Maths])\}\cup ([0,7]\times \{(R2,[u1,Math])\}\cup\dots $\newline
    	$E_M = [0,6]\times[(R1,[U1,Maths]),(R2,[U1,Maths])] \cup [1,4]\times [(R1,[U3,Maths]),(R3,[U3,Maths])] \cup \dots$}
    	\label{fig:chercheurs}
	\end{figure}

	\pimprenelle{cet exemple risque d'être retiré au profil de celui des lignes d'avion (puisqu'on a le dataset). Pas facile du tout de trouver un dataset d'espèces animales, il faudrait qu'ils portent des puces capables de nous dire quel type d'interaction ils ont eues, à quel moment ils tombent malades... ou peut être générer un graphe aléatoirement et vraisemblable ?}

@@

		We can model this system with a multilayer stream graphs. The nodes would be the individuals, and the aspect their characteristics : their species, the type of possible interaction (predator, parasite), the state of the individual (sane, infected, contagious, dead, healed ... ). The links appear when there is an interaction between two individuals, the type of the link depends on the two layers of the nodes. This system is time-dependant, and we can draw the corresponding stream.

@@

	\newpage

   	\section{Substreams and operations over substreams}

    From this definition, let us define common streams of interest. Depending on what we need, we can extract the stream of one layer, or the stream of interactions between two layers.

	We can begin to "classify" edges according to the nature of their nodes :

	\begin{defn}[Coupling, intra-layer and inter-layer edges]

@@

	\end{defn}

	\begin{figure}[h]
		\centering
		\includegraphics[width=\textwidth]{schemas/edgesCat.jpg}
		\caption{Illustration with our example of coupling edges, intra-layer edges and inter-layer edges}
		\label{exIntraInter}
	\end{figure}

	In the multilayer graph theory, the multiplex graphs have some specific properties and some tools have been created to study their properties. We define here the multiplex stream graph.

	\begin{defn}[Multiplex stream graph]	
	A {\em multiplex stream graph (MpSG)} is a graph in which edges are either coupling edges or intra-layer edges.
	\end{defn}

	\begin{defn}[Intra-layer stream graph]
	For any layer $\alpha \in L_1 \times \dots \times L_d$, the {\em intra-layer} stream graph $S^{\alpha}$ is the stream $S^{\alpha}=(T^{\alpha},V^{\alpha},W^{\alpha},E^{\alpha})$ such that $T^{\alpha} = T_{\alpha} \in T_M$ is the interval of existence of $\alpha$. $V^{\alpha}$ is the set of node-layers on the layer $\alpha$ and $W^{\alpha}$ are their time of appearance on this layer. $E^{\alpha}$ is the subset of $E_M$ containing only the links between the nodes-layers of $\alpha$.
	\end{defn}

	\begin{defn}[Inter-layer stream graph]	
	For any two layers $\alpha, \beta \in L_1\times \dots\times L_d$, the {\em inter-layer} stream graph is the stream $S^{(\alpha,\beta)} = (T, V^{\alpha,\beta},W^{\alpha,\beta},E^{\alpha,\beta})$ such that : $T^{\alpha,\beta}=T^{\alpha}\cap T^{\beta}$ is the interval of time in which $\alpha$ and $\beta$ appear simultaneously. $V^{\alpha,\beta}$ are all the node-layers of the multiplex stream graphs which are in the layers $\alpha$ and $\beta$, $W^{\alpha,\beta}$ describes theirs intervals of appearance. Then, $E^{\alpha,\beta}$ are the undirected links between nodes-layers of $\alpha$ and $\beta$ with their times of occurrence.
	\end{defn}

	\section{Intersection and union}

	After extracting some layers or inter-layer graphs, we might want to study a few layers at the same time, or to study an intersection.

	We built then the tools adapted.

	\begin{defn}[Intersection of two substreams]
	The {\em intersection} of two substreams $S_1$ and $S_2$ is defined as follows :
	\[
		S' = S_1 \cap S_2 = (T_1\cap T_2, V_1 \cap V_2, W_1 \cap W_2, E_1\cap E_2)
	\]
	\end{defn}

	\begin{defn}[Union of two substreams]
	The {\em union} of two substreams $S_1$ and $S_2$ is defined as follows :
	\[
		S' = S_1 \cup S_2 = (T', V_1 \cup V_2, W_1 \cup W_2, E' \})
	\]
	As $T'$ should be an interval, we take the shortest interval containing $T_1$ and $T_2$.
	\[
		T' = [\min(T_1,T_2),\max(T_1,T_2)]
	\]
	\[
		E' = E_1 \cup E_2 
	\]
	We note that the union stream graphs contains at least two connected components with this definition.

	Notice that the union of two sub stream graphs gives a stream graph.
	\end{defn}

	\begin{defn}[Underlying stream]
		The {\em underlying stream } $S_U(S_M)$ of $S_M$ is $(T,V_M,W_M,E_M)$. Its is the stream graph in which the nodes are the node-layers. It could be shared into clusters corresponding to the different layers. This definition is a generalisation to the one defined on \cite{mlkiv} for simple multilayer graphs.
	\end{defn}

	\begin{defn}[Aggregated stream]

		The {\em aggregated stream} $S_A(S_M)$ has the same $T$ as $S_M$. Its nodes are the nodes of $S_M$ and a link exists between two nodes if and only a link exist between two correspondent node-layers.	This definition is a generalisation to the one defined on \cite{mldd} for simple multilayer graphs.

	\end{defn}

	We can now demonstrate that the underlying stream graph is the union of all the intra and inter layer stream graphs.

	\begin{prop}
		\[
			\bigcup_{(\alpha,\beta) \in L^2} S^{(\alpha,\beta)} = S_U(M_S)
		\]
	\end{prop}
	\begin{proof}
	We call $S=\bigcup_{(\alpha,\beta) \in L^2} S^{(\alpha,\beta)}=(T^{*},V^{*},W^{*},E^{*})$ and we want to show that $S=S_U(M_S)=S_U(M) = (T_U,V_U,W_U,E_U)$.

	$T=T_U$ by definition of $S_U$. By construction of multilayer stream graphs, at each $t\in T$, at least one layer $\alpha$ is existing. Then as $T* = [\min_{\alpha \in L} (T^{\alpha}),max_{\alpha \in L} (T^{\alpha})]$, $t$ is included in $T*$. Each $T^{\alpha}$ is included in $T$, so $T^* \subseteq T$, which leads us to $T^*=T_U=T$.

	$V_M = V_U$ by definition of $V_U$.$V^{*}=\bigcup_{(\alpha,\beta) \in L^2} V^{\alpha,\beta} \subseteq V_M$. All the nodes of all the layers are included in $\bigcup_{(\alpha) \in L^2} V^{\alpha,\alpha} \subseteq V^{*}$, so we obtain the equality $V_M=V_U=V^*$.
	With the same argument we find that $W_M=W_U=W^*$.

	$E_U=E_M$ and $E^{*}=\bigcup_{\alpha,\beta \in L^2}(E^{\alpha,\beta})$ by definition. So as $E^{\alpha,\beta}$ is a subset of $E_M$ for every couple of layers $\alpha,\beta$, $E^{*} \subset E_M$. We notice too that for every $e =(t,u,\alpha,v,\beta) \in E_M$, $e$ belongs to $E^{\alpha,\beta}$. So $\mathbf{E_M=E^{*}}$.

	\end{proof}

	\section{Projections on the time}

	We can then decide to aggregate informations during the time or to select informations at a certain time.

	We can first decide to "take a picture" of the graph at a precise time t :
	\begin{defn}[Multilayer graph at time t]
   	The {\em multilayer graph at time t} $M_t$ is $M_t = (V_{M,t}, E_{M,t},V,{\cal L})$ with $V_{M,t}$ and $E_{M,t}$ containing all the node-layers and links appearing at $t$.
   	$$ V_{M,t} = \{ (u,\alpha) | (t,u,\alpha) \in W_M\} $$
   	$$ E_{M,t} = \{(u,\alpha,v,\beta) | (t,u,\alpha,v,\beta) \in E_M\}$$

   \end{defn}

	\begin{figure}[h]
		\begin{minipage}{0.49\linewidth}
			\includegraphics[width=\textwidth]{schemas/pauset.jpg}
		\end{minipage}
		\begin{minipage}{0.49\linewidth}
			\includegraphics[width=\textwidth]{schemas/pausetproj.jpg}
		\end{minipage}
		\caption{Our multilayer example at time $t=5$}
	\end{figure}

    The {\em multilayer graph induced} $M_I(S_M) = (V_{M,I}, E_{M,I}, V,L)$ of $S_M$ is the multilayer graph which gather all the layers, node-layers and links which appear during $T$.
    \begin{align*}
    	V_{M,I} = \bigcup_{t\in T} V_{M,t}\\
    	E_{M,I} = \bigcup_{t\in T} E_{M,t}\\
    \end{align*}

	We notice that many graphs are actually extracted from multilayer stream graphs, with the different operations we presented.

	For example, the page-rank algorithm uses the aggregated and induced graph of the multilayer stream graph looking like the one of our example. The temporal and multilayer informations have been "lost" in this operation but we understand well that they existed at the beginning.

@@

	We might want then to quantify the presence and the number of interactions in a multilayer stream graph. We will propose a few measures which take into account the multilayer and the stream aspect.

	The number of nodes in a graph $G=(V,E)$ is $|V|$ and the number of links is $|E|$.

	The number of nodes in a stream graph $S=(T,V,W,E)$ is $n=\frac{|W|}{|E|}=\sum_{v\in V} n_v$, $n_v$ is called the {\em contribution of v} and is equal to $\frac{|T_v|}{|T|}$.

	In multilayer graphs, the number of nodes hasn't been explicitly defined as far as we know.
	Rather than just counting the nodes or the node layers, we define the {\em number of nodes in a multilayer} as :
	\begin{align*}
		N_{nodes in multilayers}(M_l) = \frac{\text{number of node-layers}}{\text{number of layers}}
	\end{align*}	    
	Notice that if there is only one layer, we find the classical number of nodes, which means that this definition is a generalisation of the number of nodes for multilayer graphs.

	\begin{defn}[Contribution of layer]
	We define too the {\em contribution} of one layer as follows : $n_\alpha = \frac{|T_{\alpha}|}{|T|}$. The {\em number of layers} is then the sum of their contributions.
    \end{defn}

	\begin{defn}[Contribution of node-layer]
	The {\em contribution of one node-layer} in the multilayer stream graph is $n_{v,\alpha} = \frac{|T_{v,\alpha}|}{|T|}$. The {\em number of node-layers} is the sum of the contributions $N(M) = \underset{(u,\alpha)\in V_M}{\sum} n_{(u,\alpha)} = \frac{|W_M|}{|T|}$.
    \end{defn}

	Notice that if we have only one layer, we find the number of node-layers which is the number of nodes. If we take the underlying stream of the multilayer stream graph, we find too the same number of node-layers.

@@

    \begin{defn}[Contribution and number of nodes]
    The {\em the contribution} of a node describes the rate of appearance of the node on the different layers : $n_v = \frac{\sum_{\alpha \in L}|T_{v,\alpha}|}{\sum_{\alpha \in L} |T_{\alpha}|}$.

    The {\em number of nodes} is then the sum of the contribution of the nodes : 
    \begin{align}
    N_n(M) = \sum_{v\in V} n_v= \sum_{v\in V} \frac{\sum_{\alpha \in L}|T_{v,\alpha}|}{\sum_{\alpha \in L} |T_{\alpha}|} 
    \label{numberNodes}
	\end{align}

	\end{defn}

	This definition is coherent with the notions in multilayers and stream graphs : if the node appears in all the layers all the time, we find $n_v = 1$ like in multilayers. If all the times of existence are all equals, whe find $n_v=\frac{\text{number of node-layers of node v}\times |T|}{\text{number of layers}\times |T|}$. This definition is then a good generalisation of the two notions.

	\section{Uniformity}

	In stream graph the {\em uniformity between a pair of nodes} $u$ and $v$ is the ratio $\Cup U(u,v) = \frac{|T_u\cap T_v|}{|T_u \cup T_v|}$, and the graph is the sum of uniformities for all the pairs of nodes.

	The {\em uniformity } of two nodes-layers $(u,\alpha)$ and $(v,\beta)$ measures the rate of overlap presence of two node-layers:
	\begin{align*}
		\Cup((u,\alpha),(v,\beta))=\mathbb{P}( t \in T_{u,\alpha} \cap T_{v,\beta} | t \in T_{u,\alpha} \cup T_{u,\beta}) \\
		= \frac{|T_{u,\alpha}\cap T_{u,\beta}|}{|T_{u,\alpha}\cup T_{u,\beta}|}
	\end{align*}

	The {\em overlap presence} of two layers $\alpha$ and $\beta$ is defined on the same pattern :
	\begin{align*}
		\cup(\alpha,\beta) = \frac{|T_{\alpha}\cap T_{\beta}|}{|T_{\alpha}\cup T_{\beta}|}
	\end{align*}

    The {\em node-layer uniformity} measures the rate of simultaneous presences of node-layers in the multilayer-stream graph :
    \[
    	\Cup(M) = \frac{\sum_{(u,\alpha),(v,\beta) \ V_M \otimes V_M}{|T_{(u,\alpha)} \cap T_{(v,\beta)}|}}{\sum_{(u,\alpha),(v,\beta) \ V_M \otimes V_M}{|T_{(u,\alpha)}\cup T_{(v,\beta)}|}}
    \]

@@

	When $\Cup=1$, we say that the multilayer stream graph is uniform.
